---
title: "carpackage-usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{carpackage-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(carpackage)
library(tidyverse)
```

## Overview

This vignette demonstrates how to use the `carpackage` R package to clean, prepare, and model car safety data using a k-Nearest Neighbors (kNN) classifier. These steps mirror the workflow used in our DSCI 310 final project.

---

## 1. Load and Clean the Data

```{r}
# Read car dataset
df <- car_data

# Assign column names (as per UCI dataset)
colnames(df) <- c("buying", "maint", "doors", "persons", "lug_boot", "safety", "class")

# Check NAs and duplicates
df_clean <- check_data_quality(df)
head(df_clean)
```

---

## 2. Encode Ordinal Categorical Variables

```{r}
# Encode ordinal variables into numeric format
df_encoded <- df %>%
  mutate(across(-safety, encode_levels)) %>%
  mutate(safety = encode_safety(safety))
head(df_encoded)
```

---

## 3. Summarize and Visualize the Data

```{r}
# View class distribution and key feature summaries
summary_table <- summarize_data(df_encoded)
print(summary_table)
```

---

## 4. Run the kNN Classifier

```{r}
# Train/test split
df_clean <- df_encoded %>%
  na.omit() %>%
  distinct()
data_splitted <- split_data(df_clean, response_col = "safety", prop = 0.8)
knn_results <- cross_validate_knn(data_splitted$train_x, data_splitted$train_y)

# Check predictions
head(knn_results)
```

---

## 5. Evaluate Model Performance with best parameters

```{r}
# Pick best k
best_row <- knn_results[which.max(knn_results$accuracy), ]
best_k <- best_row$k

# Train final model
final_model <- train_final_model(data_splitted$train_x, data_splitted$train_y, best_k)

# Predict on test set
predictions <- predict(final_model, data_splitted$test_x)
# Calculate accuracy and view confusion matrix
performance <- compute_test_accuracy(predictions, data_splitted$test_y)
performance
```

---

## 6. Visualize the Confusion Matrix

```{r}
# Plot a heatmap-style confusion matrix
plot_conf_matrix(predictions, data_splitted$test_y)
```

---

## Conclusion

The `carpackage` provides a streamlined pipeline for preparing car evaluation data and applying a kNN classifier. With built-in support for ordinal encoding, data quality checks, and visual diagnostics, it simplifies the workflow for categorical ML tasks like car safety prediction.
